Extended Data

### 2022-03-28 doc/manuscript/Dow_manuscript_Nature.Rmd
Update Dow_manuscript_Nature.Rmd

In temperate deciduous forests, spring warming generally lengthens the period over which trees have photosynthetically active leaves[@jeong_phenology_2011;<!--@ibanez_forecasting_2010;@menzel_european_2006;@menzel_growing_1999;-->@lu_no_2022] and that over which the ecosystem is a net CO~2~ sink[@keenan_net_2014;@lu_no_2022].
Based on these observations, current terrestrial ecosystem models represent warm springs and longer growing seasons as contributing positively to annual gross primary productivity (GPP) and net CO~2~ uptake (i.e., net ecosystem exchange, NEE)[@richardson_terrestrial_2012;@ahlstrom_robustness_2012].
However, the long-term persistence of this extra C in the ecosystem, and the associated negative feedback to climate change[@keenan_net_2014], will depend on the extent to which it is allocated woody growth and therefore resides in the ecosystem for decades to centuries[@xue_global_2017;@russell_residence_2014], as opposed to being rapidly released back to the atmosphere through respiration[@jiang_fate_2020;@oishi_warmer_2018;@buermann_widespread_2018].
Model representation of C allocation to stem growth -- or woody aboveground net primary productivity (ANPP~woody~) on the ecosystem level -- does not capture known decoupling of stem growth from photosynthate production[@delpierre_temperate_2016;@delpierre_wood_2016;@fatichi_moving_2014].
As a result, the consequences of rising spring temperatures on stem growth may not be accurately represented in models[@delpierre_temperate_2016;@fatichi_moving_2014]. 
Understanding the sensitivity of ANPP~woody~ to spring temperatures is central to predicting the future of the temperate deciduous forest C sink.

@@

<!--discussion of extra C-->
Combined with widespread observations that warming springs tend to lengthen the season of C sequestration[@lu_no_2022;@finzi_carbon_2020] and increase net C sequestration[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014;@crabbe_extreme_2016;@zhou_explaining_2016;@fu_maximum_2019;@finzi_carbon_2020], our findings imply a lengthening of the period from peak stem growth to the cessation of CO~2~ uptake by the ecosystem and an increase in C allocated to functions other than stem expansion in the current or following year.
It remains theoretically possible that warm springs could augment ANPP~woody~, which, although routinely calculated based on stem growth, can be partially decoupled from it through differences in wood density or C content[@delpierre_wood_2016].
Extra C fixed in years with warm springs could potentially be allocated to formation of more C-dense wood, either through enhanced cell wall thickening (a process that lags behind stem expansion[@cuny_woody_2015]) or to a higher ratio of high-density latewood to lower-density earlywood.
However, existing evidence indicates that vessel features are most strongly controlled by summer drought stress in the previous (earlywood) or current (latewood) year, while warm springs have a neutral or negative effect on latewood width[@tardif_influence_2006;@roibu_climatic_2020;@kern_multiple_2013]<!--tardif is essential here-->.
Thus, it is unlikely that warm springs have a positive effect on total C content of annual rings or ANPP~woody~. 
The fate of any additional C fixed during years with warm springs remains unresolved, but possible destinations -- including respiration, NSC storage, and production of foliage, reproductive structures, roots[@zohner_how_2021], or root exudates -- generally have shorter residence times than woody growth.
Indeed, when GPP of a mature forest was increased through experimental CO~2~ enrichment, ANPP~woody~ remained unchanged, while additional C was released back to the atmosphere on relatively short time scales through enhanced respiration[@jiang_fate_2020].
Consistent with this, it has been observed that C gains from an earlier spring can be offset through fall or winter respiration[@oishi_warmer_2018;@buermann_widespread_2018], although even the C in shorter-lived pools would often be carried over into the following year[@trumbore_quantifying_2002].
Thus, observed augmentation of NEE by warm springs[@zhou_explaining_2016;@fu_maximum_2019;@keenan_net_2014] is likely to be compensated by increased respiration in subsequent years.

>>>>

In temperate deciduous forests, spring warming generally lengthens the period over which trees have photosynthetically active leaves[@jeong_phenology_2011;<!--@ibanez_forecasting_2010;@menzel_european_2006;@menzel_growing_1999;-->@lu_no_2022] and that over which the ecosystem is a net CO~2~ sink[@keenan_net_2014;@lu_no_2022].
Based on these observations, current terrestrial ecosystem models represent warm springs and longer growing seasons as contributing positively to annual gross primary productivity (GPP) and net CO~2~ uptake (i.e., net ecosystem exchange, NEE)[@richardson_terrestrial_2012;@ahlstrom_robustness_2012;@buermann_widespread_2018].
However, the long-term persistence of this extra C in the ecosystem, and the associated negative feedback to climate change[@keenan_net_2014], will depend on the extent to which it is allocated woody growth and therefore resides in the ecosystem for decades to centuries[@xue_global_2017;@russell_residence_2014], as opposed to being rapidly released back to the atmosphere through respiration[@jiang_fate_2020;@oishi_warmer_2018].
Model representation of C allocation to stem growth -- or woody aboveground net primary productivity (ANPP~woody~) on the ecosystem level -- does not capture known decoupling of stem growth from photosynthate production[@delpierre_temperate_2016;@delpierre_wood_2016;@fatichi_moving_2014].
As a result, the consequences of rising spring temperatures on stem growth may not be accurately represented in models[@delpierre_temperate_2016;@fatichi_moving_2014]. 
Understanding the sensitivity of ANPP~woody~ to spring temperatures is central to predicting the future of the temperate deciduous forest C sink.

@@

<!--discussion of extra C-->
Combined with widespread observations that warming springs tend to lengthen the season of C sequestration[@lu_no_2022;@finzi_carbon_2020] and increase net C sequestration[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014;@crabbe_extreme_2016;@zhou_explaining_2016;@fu_maximum_2019;@finzi_carbon_2020], our findings imply a lengthening of the period from peak stem growth to the cessation of CO~2~ uptake by the ecosystem and an increase in C allocated to functions other than stem expansion in the current or following year.
It remains theoretically possible that warm springs could augment ANPP~woody~, which, although routinely calculated based on stem growth, can be partially decoupled from it through differences in wood density or C content[@delpierre_wood_2016].
Extra C fixed in years with warm springs could potentially be allocated to formation of more C-dense wood, either through enhanced cell wall thickening (a process that lags behind stem expansion[@cuny_woody_2015]) or to a higher ratio of high-density latewood to lower-density earlywood.
However, existing evidence indicates that vessel features are most strongly controlled by summer drought stress in the previous (earlywood) or current (latewood) year, while warm springs have a neutral or negative effect on latewood width[@tardif_influence_2006;@roibu_climatic_2020;@kern_multiple_2013]<!--tardif is essential here-->.
Thus, it is unlikely that warm springs have a positive effect on total C content of annual rings or ANPP~woody~. 
The fate of any additional C fixed during years with warm springs remains unresolved, but possible destinations -- including respiration, NSC storage, and production of foliage, reproductive structures, roots[@zohner_how_2021], or root exudates -- generally have shorter residence times than woody growth.
Indeed, when GPP of a mature forest was increased through experimental CO~2~ enrichment, ANPP~woody~ remained unchanged, while additional C was released back to the atmosphere on relatively short time scales through enhanced respiration[@jiang_fate_2020].
Consistent with this, it has been observed that C gains from an earlier spring can be offset through fall or winter respiration[@oishi_warmer_2018], although even the C in shorter-lived pools would often be carried over into the following year[@trumbore_quantifying_2002].
Thus, observed augmentation of NEE by warm springs[@zhou_explaining_2016;@fu_maximum_2019;@keenan_net_2014] is likely to be compensated by increased respiration in subsequent years.
